---
title: "Micro Capstone"
author: "Scott Waugh"
output:
  html_document:
    df_print: paged
---

Read in excel sheet with data from baseball reference from 1969 to 2021 with about 162 games (excluding 1981, 1994, 1995, and 2020) and removing random rows that show up

```{r}
baseball = read.table("BaseballCapstoneData.csv", header = TRUE, sep=",")
baseball$X = NULL
baseball$X.1 = NULL
baseball$X.2 = NULL
baseball$X.3 = NULL
baseball$X.4 = NULL
baseball$X.5 = NULL
```

 Run the full model looking at response variable y with all the variables
 
```{r}
startmodel = lm(W ~ BatAge + RG + R + H + SB + TB + HR + RBI + SB + CS + BB + SO + BA + OBP + SLG + OPS + PAge + RAG + ERA + CG + tSho+ cSho + SV + HA + RA + ER + HRA + BBA + SOA + FIP + WHIP, data = baseball)

summary(startmodel)
```
 
 Run stepwise to find best model
 
```{r}
require(MASS)
min_model = lm(W ~ 1, data = baseball)
max_model = formula(W ~ BatAge + RG + R + H + SB + TB + HR + RBI + SB + CS + BB + SO + BA + OBP + SLG + OPS + PAge + RAG + ERA + CG + tSho+ cSho + SV + HA + RA + ER + HRA + BBA + SOA + FIP + WHIP, data = baseball)
best_model = step(min_model, direction = "both", scope = max_model)
```
 
 run new stepwise model
 
```{r}
stepwiseModel = lm(W ~ WHIP + R + RAG + SV + CG + tSho + OBP + BatAge + HA + BBA + 
    HR + H + BA + CS + SO + BB + FIP, data = baseball)

summary(stepwiseModel)
```
 
 run test for collinearity 
 
```{r}
require(regclass)
stepwiseModel = lm(W ~ WHIP + R + RAG + SV + CG + tSho + OBP + BatAge + HA + BBA + 
    HR + H + BA + CS + SO + BB + FIP, data = baseball)
VIF(stepwiseModel)
```
 
make small data frame for correlation matrix

```{r}
baseballNew <- baseball[, colnames(baseball)[c(3,7,9,10,13,16,17,18,19,20,24,26,27,29,30,34,36,37)]]
```
 
run correlation matrix
 
```{r}
round(cor(baseballNew),3)
```

Model after removing collinerarity 

```{r}
require(regclass)
modelfinal = lm(W ~ WHIP + SV + tSho + OBP + BatAge + 
    HR + CS + SO + BB + FIP, data = baseball)
VIF(modelfinal)
summary(modelfinal)
```
remove batters age since not significant

```{r}
modelfinal = lm(W ~ WHIP + SV + tSho + OBP + 
    HR + CS + SO + BB + FIP, data = baseball)
summary(modelfinal)
```

predict wins with formula

```{r}
baseball$Pred = predict(modelfinal, baseball)
```

 calculate residuals and plot
 
```{r}
baseball$Residuals = residuals(modelfinal)
plot(baseball$W, baseball$Residuals, xlab = "Wins", ylab = "Residual")  
abline(a=0, b=0)
hist(baseball$Residuals, xlab = "Residuals")
qqnorm(baseball$Residuals)
qqline(baseball$Residuals)
```
 Start over using 2011-2021 excluding 2020 and see if there is any difference 
 
```{r, echo=FALSE}
baseball10years = read.table("BaseballCapstoneData10Years.csv", header = TRUE, sep=",")
baseball10years$X = NULL
baseball10years$X.1 = NULL
baseball10years$X.2 = NULL
baseball10years$X.3 = NULL
baseball10years$X.4 = NULL
baseball10years$X.5 = NULL
```

 Run the full model looking at response variable y with all the variables
 
```{r, echo=FALSE}
startmodel = lm(W ~ BatAge + RG + R + H + SB + TB + HR + RBI + SB + CS + BB + SO + BA + OBP + SLG + OPS + PAge + RAG + ERA + CG + tSho+ cSho + SV + HA + RA + ER + HRA + BBA + SOA + FIP + WHIP, data = baseball10years)

summary(startmodel)
```
 
 Run stepwise to find best model
 
```{r}
require(MASS)
min_model = lm(W ~ 1, data = baseball10years)
max_model = formula(W ~ BatAge + RG + R + H + SB + TB + HR + RBI + SB + CS + BB + SO + BA + OBP + SLG + OPS + PAge + RAG + ERA + CG + tSho+ cSho + SV + HA + RA + ER + HRA + BBA + SOA + FIP + WHIP, data = baseball10years)
best_model = step(min_model, direction = "both", scope = max_model)
```
 
run new stepwise model and check for collinerarity
 
```{r}
require(regclass)
stepwiseModel = lm(W ~ WHIP + R + RAG + SV + tSho + CG + BBA + HA + CS + BB + H + 
    BA + HRA + RBI, data = baseball10years)

summary(stepwiseModel)
VIF(stepwiseModel)
```

make small data frame for correlation matrix

```{r}
baseball10New <- baseball10years[, colnames(baseball10years)[c(3,9,10,14,16,17,19,24,26,27,29,30,33,34,37)]] 
```
 
run correlation matrix
 
```{r}
round(cor(baseball10New),3)
```

Model after removing collinerarity

```{r}
require(regclass)
modelfinal10 = lm(W ~ WHIP + SV + tSho + CG + BBA + CS + 
    BA + HRA + RBI, data = baseball10years)
VIF(modelfinal10)
summary(modelfinal10)
```

remove CS and BA since not significant

```{r}
modelfinal10 = lm(W ~ WHIP + SV + tSho + CG + BBA + HRA + RBI, data = baseball10years)
summary(modelfinal10)
```

predict wins with formula

```{r}
baseball10years$Pred = predict(modelfinal10, baseball10years)
```

calculate residuals and plot
 
```{r}
baseball10years$Residuals = residuals(modelfinal10)
plot(baseball10years$W, baseball10years$Residuals, xlab = "Wins", ylab = "Residual")  
abline(a=0, b=0)
hist(baseball10years$Residuals, xlab = "Residuals")
qqnorm(baseball10years$Residuals)
qqline(baseball10years$Residuals)
```


